Molecular dynamics simulation of synchronization in driven particles

### 2021-01-16 manuscript/manuscript_ajp_2020.tex
updated fig. 1 to have good colors and arrow lengths

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  %Simplified models
  %such as 
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  move in 
  synchronized patterns,
  as observed in 
  experiments %al systems 
  of %interest to condensed matter physics %including 
  magnetically driven colloidal particles
  confined by light-fields.
  %exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to external stimulus \cite{Pikovsky2003}.
Many everyday systems
exhibit in-phase oscillations of components 
such as the
flickering patterns of
candle flames coupled by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses \cite{Arane2009}, 
and metronomes coupled through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization -- %such as 
birds coordinate wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternate croaking patterns \cite{Aihara2014},
humans clap in time with music \cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.
External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance,
an electrical pacemaker %(an electrical generator)
pulses to regulate a heart beat,
and 
a flashing light can modify the
pattern of flashing fireflies .% \cite{Agrawal2013}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
%two periodic oscillators in close proximity
to swing in time %, i.e. to synchronize,
due to interactions through the wall,  
no matter the original phase of the clocks \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms adjust %couple
to a frequency ratio or mode 
of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
Synchronization 
is often studied with simple computational models
and can be visualized 
with phase
plots or Lissajous figures.
In phase space 
a mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
%({\it As we will demonstrate in phase plots in Figs. X, Y, Z})
It should be noted that oscillation phase and phase space
  refer to different uses of the word ``phase.''  Oscillation
  phase $\phi_0$ is the constant portion of the argument in a periodic function
  $\sin{(\omega t + \phi_0)}$ while phase space refers
  to plots of velocity versus position (such as $v_y$ vs. $y$)
  that are used to characterize oscillating systems.

@@

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions \cite{Ashkin1997}.
A circular optical trap is designed so the center is
a local energy minimum where 
a colloid in the beam center is subject to a
uniform distribution of photon bombardment. %, making 
Colloids at a trap edge
experience a net force
due to the uneven 
distribution of photon collisions on its surface.
Depending on the shape of the trap and location of the particle in the trap,
the radiation pressure either pushes colloids toward the local minima
or ejects it from the trap.
A single optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

Experiments of colloid particles confined
in optical traps arrays or patterns 
subject to external driving forces 
have
been used to examine the microscopic dynamics
of individual particles during mode-locking \cite{Juniper2015}.
This produced step-by-step diagrams
of particle dynamics that cannot be 
imaged in electronic systems such as Josephson junctions.
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Thus the study of
mode-locking is a useful problem
of complex quantum mechanical systems.

@@

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The landscape is modulated in the y-direction
with the periodic function 
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the minima
 with simulation units of energy $E_0 = 1$. 
 %We define the substrate strength to be
 %the maximum force the substrate can exert on a particle 
 %$A_p = 2\pi U_0/L$.
 %
 In experiment 
 complex environments can be
 engineered with light by %, typically made 
 using diffraction gratings to create a patterned
 potential energy landscape 
 with one or more laser beams \cite{}.
 %
 We plot this function in 
 Fig.~\ref{fig:landscape0}
 we use $N_p = 3$ to illustrate the features of
 Eq.~\ref{eq:ysubstrate}.
 In Fig.~\ref{fig:landscape0}(a) we show 
 the $x-y$ plane with a contour plot
 of the the 2D potential energy landscape,
 where 
 the maxima are colored red and the minima colored blue.
 The confining force on a particle $i$
 %arises from a potential function 
 %creating a landscape of potential minima and maxima
 %that modify the force on a particle as a function of position
 is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 In Fig.~\ref{fig:landscape0}(b) we plot the function
 $U(y)$ to illustrate how the magnitude
 $\vec{F}^{l}_i$ is calculated from particle position $y_i$.  
 %in the $y-$direction.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  The period of the landscape, $\lambda = L/N_p$ 
  where $N_p = 3$, is marked in red.
  The force due to the landscape %$\vec{F}^l$
  is calculated from the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^d$
  is parallel to the y-axis.
  The landscape is shown with a contour plot,
  with maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is shown in the region between minima and maxima
  subject to competing forces due to the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$,
  The particle in (a) is shown at the same $y-$position,
  where the magnitude of force $\vec{F}^l$ is calculated as the slope of this function,
  directed along the $y-$direction in panel (a).
  }
\label{fig:landscape0}
\end{figure}
\end{center}

@@

\begin{table}[h!]
\centering
\caption{Simulation units.}
\begin{ruledtabular}
\begin{tabular}{c c p{5cm}}
Quantity & Unit \\
\hline
length &  $a_0 = 1$ \\
energy & $E_0 = 1$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $F_0 = E_0 / a_0$\\
velocity &  $ a_0 / \tau $ \\
viscosity coefficient & $\eta = F_0 \tau / a_0 = 1 $\\
time &  $ \tau = \eta a_0 / F_0 $ \\

@@

In Fig.~\ref{fig:0} 
the driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$ time units.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^d(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
(i.e. $\sin{(\omega t)} = 1$).

@@

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$, the spatial period of the potential
to show how the particle moves between substrate minima.  
The initial particle position is $y = 1$,
where the $F^l$ is negative, 
causing the particle to move to
a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
becomes locked to the driving period $T$.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ in one time period $\Delta t = T$,
with 
the average velocity 
$\bar{v}_y = 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$ 
(i.e. $\sin{(\pi y / \lambda)} = 1$).
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
Once the particle motion
coincides with the applied drive, 
the particle position $y$
crosses the intersection of
the horizontal and vertical lines.
This indicates 
the driving force is maximum positive when the
force on the particle from the landscape is maximum negative,
indicating synchronization.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $. 
We invite interested students to repeat this
simulation for any initial particle position %of $y$
and a range of simulation constants
to demonstrate that synchronization in this model
does not depend on a particular set of conditions.
%but is a direct consequence of the simplicity of the model.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion.
If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima
with no net velocity.  
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1

The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and length of the particle motions 
forward and backward in the landscape.
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
We explore this in Ex.~\ref{ex:backwards}.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $\Delta t = 10^5$ simulation timesteps
and 
sample the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.
We plot the results in Fig.~\ref{fig:1}
for $F^{ac} = 0.05$, $f=0.01$, and the landscape force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
The steps in $\langle v_y \rangle$ occur because 
the average velocity of the particle
is constant across a range of $F^{dc}$.
This synchronization
of the particle velocity can be measured
in terms of the landscape period.
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle.
The average velocity $\langle{v}_y\rangle$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
At low $F^{dc}$, the particle starts and ends in the same minima.

@@

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.

@@

We perform this sweep for several values
of $F^{ac}$ including the limiting case that the applied force
is a constant value.
When $F^{ac} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{dc}$ curve.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$.
  We fix the remaining parameters in $F^d(t)$ as $F^{ac}=0.05$ and $f = 0.01$ as in Fig.~\ref{fig:0}}.
\label{fig:1}
\end{figure}
\end{center}

@@

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.

>>>>

\begin{abstract}
  Synchronization
  plays a key role in many physical processes.
  %Simplified models
  %such as 
  %What problem did you study and why is it important?
  Modeling synchronization
  with simulations and table-top experiments
  can 
  provide insight
  to complex behaviors
  in the natural world.
  Driving particles
  across a washboard potential energy landscape
  is a 
  common 
  technique to isolate complex 
  synchronized patterns 
  in both simulations and experiments.
  %as observed in 
  %experiments %al systems 
  %of %interest to condensed matter physics %including 
  %magnetically driven colloidal particles
  %confined by light-fields.
  %exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model %that
  %exhibits
  %provides insight
  %to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We demonstrate how to 
  %visualize
  %the system with
  %numerical representations of the landscape
  %and
  %animations of the particle motion.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

\section{Introduction} %
%
%Outline
%------------------------------------------------------------------
%synchronization = widest audience possible
Synchronization is a universal phenomena
in which individual oscillators adjust frequency due
to external stimulus \cite{Pikovsky2003}.
Many everyday systems
exhibit in-phase coupled oscillations %of components 
such as the
flickering patterns of
candle flames mediated by temperature fluctuations \cite{Okamoto2016},
vibrations of singing wineglasses interacting 
through sound waves \cite{Arane2009}, 
and metronomes vibrating through a supporting platform \cite{Jia2015}.
Biological systems benefit from cooperative
synchronization -- %such as 
birds coordinate wing flaps
to optimize energy use during flight \cite{Portugal2014},
frogs alternate croaking patterns \cite{Aihara2014},
humans clap in time with music \cite{Tranchant2016}
and 
at a cellular level, 
neurons simultaneously fire in cardiac muscle \cite{MartinHall1999}
and brain tissue \cite{Singer1999}.
External forcing can cause or regulate 
synchronization. % is often achieved via external forcing
%such as when
%single particle oscillators
For instance,
an electrical pacemaker %(an electrical generator)
pulses to regulate a heart beat,
and 
a flashing light can modify the
pattern of flashing fireflies .% \cite{Agrawal2013}.

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
A particular form of 
synchronization is phase-locking or mode-locking,
which first appeared in the scientific literature
with 
Huygens' 1665 experiments on
the 
motions of two wall-mounted pendulum clocks.
Huygens demonstrated the tendency of
the pendula 
to swing in time, 
even when started out of phase,
due to interactions through the wall \cite{Bennett2002}.
%phase locking
Dynamical systems %with multiple frequencies
exhibit phase-locking 
when oscillators with different rhythms adjust %couple
to a frequency ratio or mode 
of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
Synchronization 
is often studied with simple computational models
and can be visualized 
with phase
plots or Lissajous figures.
In phase space 
a mode-locked system is confined to 
a closed loop in a 
parametric space showing the relationship between two periodic functions
with patterns determined by the mode, % frequency ratio,
first reported in 1857 \cite{Lissajous1857}
and easily generated with a computer code 
or oscilloscope \cite{Tong1997}. 
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
%({\it As we will demonstrate in phase plots in Figs. X, Y, Z})
It should be noted that the term phase
has two scientific meanings, 
``oscillation phase'' and ``phase space''
%refer to different uses of the word ``phase.''
Oscillation phase $\phi_0$ is the constant portion of the argument in a periodic function
  $\sin{(\omega t + \phi_0)}$ while phase space refers
  to plots of velocity versus position (such as $v_y$ vs. $y$)
  that are used to characterize oscillating systems.

@@

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions \cite{Ashkin1997}.
A circular optical trap is designed
with a local energy minimum 
at its center.
A colloid in the beam center is subject to a
uniform distribution of photon bombardment,
and motions away from the center create a net force
due to an uneven 
distribution of photon collisions on its surface.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either restores colloids toward the local minima
or ejects it from the trap.
A single optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

In experiments  
 complex environments can be
 engineered with light by %, typically made 
 using diffraction gratings to create a patterned
 potential energy landscape 
 with one or more laser beams \cite{}.
Experiments of colloid particles confined
in optical traps arrays or patterns 
subject to external driving forces 
have
been used to examine the microscopic dynamics
of individual particles during mode-locking \cite{Juniper2015}.
This produced step-by-step diagrams
of particle dynamics that cannot be 
imaged in electronic systems such as Josephson junctions.
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Thus the study of
mode-locking is a useful problem
of complex quantum mechanical systems.

@@

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The landscape is modulated in the y-direction
with the periodic function 
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the minima
 with simulation units of energy $E_0 = 1$. 
 %We define the substrate strength to be
 %the maximum force the substrate can exert on a particle 
 %$A_p = 2\pi U_0/L$.
 %
 We plot this function in 
 Fig.~\ref{fig:landscape0}
 for $N_p = 3$. %to illustrate the features of
 In Fig.~\ref{fig:landscape0}(a) we show 
 the $x-y$ plane with a contour plot of $U(y)$ %Eq.~\ref{eq:ysubstrate}
 to illustrate
 the 2D potential energy landscape,
 where 
 the maxima are colored red and the minima colored blue.
 The confining force on a particle $i$
 %arises from a potential function 
 %creating a landscape of potential minima and maxima
 %that modify the force on a particle as a function of position
 is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 In Fig.~\ref{fig:landscape0}(b) we plot the function
 $U(y)$ to illustrate how the magnitude
 $\vec{F}^{l}_i$ is calculated from particle position $y_i$.  
 %in the $y-$direction.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  The period of the landscape is $\lambda = L/N_p$,  
  %is marked in red,
  where $N_p = 3$.  
  The force due to the landscape %$\vec{F}^l$
  is calculated from the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^d$
  is parallel to the y-axis.
  The landscape is shown with a contour plot,
  with maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is shown in the region between minima and maxima
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$.  
  The particle in (a) is shown at the same $y-$position.
  The slope of $U_l(y)$ is the 
  magnitude of force $\vec{F}^l$. %is calculated
  %as the slope of this function,
  %directed along the $y-$direction in panel (a).
  }
\label{fig:landscape0}
\end{figure}
\end{center}

@@

\begin{table}[h!]
\centering
\caption{Simulation units and approximate values
  in experiment.}
\begin{ruledtabular}
\begin{tabular}{c c c } %p{5cm}}
Quantity & Simulation Unit & Experimental range\\
\hline
length &  $a_0 = 1$ & $ < a_0 <$\\
energy & $E_0 = 1$ & $ < E_0 <$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $F_0 = E_0 / a_0$ & $<F_0<$\\
velocity &  $ a_0 / \tau $ &  $ < v < $\\
viscosity coefficient & $\eta = F_0 \tau / a_0 $ & $ <\eta<$\\
time &  $ \tau = \eta a_0 / F_0 $ & $ < \tau <$\\

@@

In Fig.~\ref{fig:0} 
the driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$ time units.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^d(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
-- i.e. $\sin{(\omega t)} = 1$.

@@

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$, the spatial period of the potential
to show how the particle moves between substrate minima.  
The initial particle position is $y = 1$,
where the $F^l$ is negative, 
causing the particle to move to
a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
becomes locked to the driving period $T$.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ in one time period $\Delta t = T$,
with 
the average velocity 
$\bar{v}_y = 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
Once the particle motion
coincides with the applied drive, 
the particle position $y$
crosses the intersection of
the horizontal and vertical lines.
This indicates 
the driving force is maximum positive when the
force on the particle from the landscape is maximum negative,
indicating synchronization.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $. 
We invite interested students to repeat this
simulation for any initial particle position %of $y$
and a range of simulation constants
to demonstrate that synchronization in this model
does not depend on a particular set of conditions.
%but is a direct consequence of the simplicity of the model.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$,
as shown in Fig.~\ref{fig:1}(b).
We keep $F^{dc}$ constant
for $\Delta t = 10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$.
We plot the results in Fig.~\ref{fig:1}
for $F^{ac} = 0.05$, $f=0.01$, and the landscape force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
The steps in $\langle v_y \rangle$ occur because 
the average velocity of the particle
is constant across a range of $F^{dc}$.
This synchronization
of the particle velocity can be measured
in terms of the landscape period.
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle.
The average velocity $\langle{v}_y\rangle$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.

If $A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima
with no net velocity.
In Fig.~\ref{fig:1}, this occurs at low
low $F^{dc}$ when $\langle v_y \rangle$ is zero.

@@

We explore a range of parameters in Ex.~\ref{ex:parameters}
where 
we perform this sweep for several values
of $F^{ac}$ including the limiting case that the applied force
is a constant value.
When $F^{ac} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{dc}$ curve. 
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$,
  the constant parameter of $F^d(t)$ defined in Eq.~\ref{eq:drive}.
  The remaining parameters in $F^d(t)$ are as in Fig.~\ref{fig:0}, $F^{ac}=0.05$ and $f = 0.01$.}
\label{fig:1}
\end{figure}
\end{center}

@@

\item
  \label{ex:parameters}

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.
